---
title: "Group Project #2"
author: "Joshua Gray, Vaijayanti Deshmukh, & Marissa Mowers"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: 3
---


```{r, echo=FALSE, result='hide'}
library(kableExtra)
library(gssr)
library(gssrdoc)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(ezids)
library(RColorBrewer)
library(psych)
library(broom)
library(dplyr)
library(pROC)
library(caret)
library(rattle)
library(rsample)
library(gssr)
library(gssrdoc)
library(rpart)
library(rattle)
library(rsample)
library(caret)
library(RColorBrewer)
library(psych)
library(broom)
library(dplyr)
library(pROC)
library(caret)

file_path <- "C:\\Users\\mnmow\\OneDrive\\Desktop\\GWU\\Fall 2024\\DATS6101 Intro\\DATS-6101-Project\\Project 2\\GSS_data.csv"
gss <- read.csv(file_path)

str(gss)

gss <- na.omit(gss)

gss$region <- factor(gss$region, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9), labels = c("new england", "middle atlantic", "east north central", "west north central", "south atlantic", "east south atlantic", "west south central", "mountain", "pacific"))
gss$sex <- factor(gss$sex, levels = c(0, 1), labels = c("Male", "Female"))
gss$race <- factor(gss$race, levels = c(1, 2, 3), labels = c("White", "Black", "Other"))
gss$relig <- factor(gss$relig, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13), labels = c("Prot", "Cath", "Jew", "None", "Other", "Budd", "Hind", "East", "Islam", "Orth", "Christ","Nativ", "Nonden"))


str(gss)
```
# Introduction

  Our research project explores predictors of confidence in the executive branch of the federal government (confed) among U.S. adults, focusing on identifying the variables that most effectively predict confidence in governmental institutions. Specifically, we address the question: Which variable or combination of variables best predicts confidence in the executive branch of the U.S. government, and how has this confidence changed over time across different demographic groups? To answer this, we analyze key demographic variables such as age, income, and education alongside race and political views (polviews), both with and without interaction terms reflecting the political party in executive office during each survey year. This approach allows us to capture how demographic and contextual factors interact to shape confidence in federal institutions over time.

  We include polviews to complement confed in order to capture both institutional trust and personal political identity. While confed assesses trust in governmental institutions, which is shaped by broader political and economic contexts, polviews provides insight into individual ideological leanings. Examining these variables together enables us to uncover the interplay between trust in federal institutions and personal political identity, particularly during pivotal moments in U.S. history. This dual analysis helps us determine whether and how shifts in institutional trust align with or diverge from ideological trends over time, offering a richer understanding of the factors influencing political confidence.

  To conduct this analysis, we use the General Social Survey (GSS) dataset, a widely used resource for social science research conducted by the National Opinion Research Center (NORC) at the University of Chicago. The GSS provides representative data on U.S. adults, collected through random sampling since 1972. For our study, we create a subset of the data, focusing on surveys from 1974 to 1980 and 2016 to 2022, encompassing 22,399 observations. These periods are chosen to highlight key moments in U.S. history, including the aftermath of the Watergate scandal, economic challenges of the 1970s, and the politically polarized environment of the late 2010s and early 2020s. Through comparing these timeframes, we identify parallels between historical and recent sentiments during times of conflict, providing insights into how public trust in government evolves across generations.


# Data Analysis

  We added polviews to complement confed (confidence in the federal branch) to capture both institutional trust and personal political identity. Confed assesses trust in governmental institutions, which is often shaped by broader political and economic contexts, while polviews provides insight into individual ideological leanings. By examining both variables, we can better understand the interplay between trust in federal institutions and individual political identity, especially during pivotal moments in U.S. history. This dual analysis allows us to explore whether and how shifts in institutional trust align with or diverge from ideological trends over time.

## EDA

The summary statistics provide key insights into the dataset, particularly regarding political alignment, time periods, and demographic characteristics. The variables confed and polviews are binary, with 0 representing Republicans and 1 representing Democrats. The mean of polviews (0.4754) suggests a fairly balanced sample, with a slight skew toward Republican views. Similarly, the variable confed, which measures confidence in the federal branch, shows a mean of 0.6252, indicating that most respondents lack trust in the federal government. The variable era, coded as 0 for earlier years (1977, 1978, 1980, and 1982) and 1 for more recent years (2018, 2021, and 2022), has a mean of 0.5417, reflecting a slightly higher representation of responses from modern years. The pres variable (0 for Republican presidencies and 1 for Democratic presidencies) has a mean of 0.5582, suggesting that more responses were collected during Democratic presidencies. Similarly, incumbency, which represents whether the current president was an incumbent, has a mean of 0.7494, indicating that a significant portion of respondents were surveyed during periods of incumbency.

Demographic variables provide additional context for the analysis. The age of respondents ranges from 20 to 89 years, with a median of 52, reflecting an older-skewing sample. The income variable, categorized from 1 (low) to 5 (high), shows a median of 3 and a mean of 3.21, suggesting that most respondents fall into middle-income categories. Education (educ), measured in years, has a median of 14 and a mean of 14.02, indicating a well-educated sample overall. These demographic and temporal variables provide a solid foundation for understanding the interaction between political views, institutional trust, and broader socio-economic trends across distinct historical periods. The balanced representation of earlier and modern eras ensures that changes in these dynamics can be effectively explored.

```{r, echo=TRUE}
library(knitr)
library(dplyr)
library(ezids)

xkablesummary(data.frame(confed = gss$confed))
xkablesummary(data.frame(polviews = gss$polviews))
xkablesummary(data.frame(age = gss$age))
xkablesummary(data.frame(income = gss$income))
xkablesummary(data.frame(educ = gss$educ))
xkablesummary(data.frame(era = gss$era))
xkablesummary(data.frame(pres = gss$pres))
xkablesummary(data.frame(incumbency = gss$incumbency))
```


## Polviews bivariate graph

This bar chart illustrates the frequency of political views, categorized as Republican (red) and Democrat (blue), across select years: 1977, 1978, 1980, 1982, 2018, 2021, and 2022. These years were chosen to capture critical periods of political and societal shifts in U.S. history. The late 1970s and early 1980s coincide with significant political changes, including the aftermath of the Watergate scandal, economic challenges like stagflation, and the rise of conservatism with the election of Ronald Reagan in 1980. These factors likely influenced public trust in government and ideological realignments. Including the early 1980s helps capture the consolidation of Republican dominance during this transformative era.The later years (2018–2022) focus on contemporary political polarization and major events such as the Trump presidency, the COVID-19 pandemic, the Black Lives Matter movement, and the 2020 presidential election, which brought heightened political engagement and shifts in partisan identity. These years also reflect the growing divide in institutional trust and political alignment during periods of intense national debate.


```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(polviews))) +
  geom_bar(position = "dodge") + 
  scale_fill_manual(
    values = c("0" = "red", "1" = "blue"), 
    labels = c("Republican", "Democrat"), 
    name = "Political View"
  ) +
  labs(
    title = "Frequency of Political Views Over the Years",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Confed bivariate graph

In comparison, this bar chart depicts the frequency of confidence in the executive branch of the federal government across various years, with "Trust" shown in green and "No Trust" in purple. The data reveals significant shifts in public confidence over time. In earlier years such as 1977 and 1978, the majority of respondents reported "No Trust," with relatively low levels of trust in the executive branch. Confidence in the executive branch appears to decrease further by 1980 and 1982, corresponding to key political and economic challenges during this period, such as inflation and dissatisfaction with federal leadership. By 2021 and 2022, confidence levels increase, as indicated by the green bars approaching or surpassing previous lows, reflecting heightened political engagement during a period of political polarization and major national events. These trends provide insight into how public trust in the executive branch responds to broader historical and political contexts over time.


```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(confed))) +
  geom_bar(position = "dodge") + 
  scale_fill_manual(
    values = c("0" = "lightgreen", "1" = "orchid"), 
    labels = c("Trust", "No Trust"), 
    name = "Confidence in the Executive Branch"
  ) +
  labs(
    title = "Frequency of Confidence in the Executive Branch Over the Years",
    x = "Years",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Who R Voted for Bivariate Graph

The clustered bar chart illustrates voting preferences by year, comparing support for Republican (red) and Democrat (blue) candidates over different time periods. Each cluster represents a specific year, with the bars showing the count of individuals who voted for each party. The chart highlights trends in political alignment over time, with notable shifts in the relative support for each party. For example, during the earlier years (1977–1982), the distribution of votes appears more balanced, while in recent years (2021–2022), there is a discernible increase in Democratic votes compared to Republican votes. This visualization offers insights into how voting preferences have evolved across decades and provides a foundation for exploring potential factors influencing these patterns, such as changes in political, social, or economic contexts.

```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(pres))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(
    values = c("0" = "red", "1" = "blue"),
    labels = c("Republican", "Democrat"),
    name = "Pres"
  ) +
  labs(
    title = "Clustered Bar Chart of Voting Preferences by Year",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Tetrachoric Correlation 

This table presents the tetrachoric correlation matrix for three variables: the respondent's political views (polviews), who they voted for (pres), and whether the incumbent party held the presidency during the survey year (incumbency). The high correlation between polviews and pres (0.7749) indicates that political views strongly influence voting behavior, creating redundancy in the analysis. Including both variables risks multicollinearity, which can obscure the independent effects of other predictors. Additionally, incumbency has a weak correlation with both polviews (0.0164) and pres (-0.0566), suggesting it captures different, less impactful information. However, incumbency is contextually tied to the survey year, which introduces external bias. These correlations highlight why analyzing these variables together is problematic, as their statistical and conceptual overlap complicates efforts to disentangle their unique contributions, leading to potentially biased or redundant results.

```{r, echo=FALSE}


binary_vars <- gss %>% select(polviews, pres, incumbency)

binary_vars <- binary_vars %>% mutate_all(as.numeric) %>% na.omit()
lapply(binary_vars, unique)

tetra_corr <- psych::tetrachoric(binary_vars)
print(tetra_corr$rho)

```


# Confidence in the Federal Government Trees

When it comes to comparing time intervals, we wanted to see if there was a noticable difference in predictability. In doing so, we could gain some insight on how diversity of thought has changed among the american people. An increase in accuracty could indicate lower diversity of thought among identity groups, and a decrease in accuracy could indicate an increase in diversity of thought among the same groups. To do this, we used classification trees to see if there was a difference in preformance between our two time intervals. 

Here is a look at how our first pair of models preformed with confed as the response variable. 


```{r}
gss$polviews <- factor(gss$polviews, levels = c(0, 1), labels = c("Cons", "Lib"))
gss$confed <- factor(gss$confed, levels = c(0, 1), labels = c("Trust", "No Trust"))
gss$pres <- factor(gss$pres, levels = c(0, 1), labels = c("Rep", "Dem"))


# Split training and test data sets
gss_split <- initial_split(gss[gss$era == 0,], prop = 0.70, strata = confed)
gss_test <- testing(gss_split)
gss_train <- training(gss_split)

#Build Tree using Political Categories
set.seed(1)
gssfit <- rpart(confed ~ sex + age + race + relig + region + income + educ + incumbency + pres + polviews, data=gss_train, method="class")
summary(gssfit)
#fancyRpartPlot(gssfit, main = "Prior Era Classification Tree for Confed")
# Testing and Validation
pred <- predict(gssfit, gss_test, type = "class")
mat <- confusionMatrix(gss_test$confed, pred)
mat
```

```{r}
# Split training and test data sets
gss_split <- initial_split(gss[gss$era == 1,], prop = 0.70, strata = confed)
gss_test <- testing(gss_split)
gss_train <- training(gss_split)

#Build Tree using Political Categories
set.seed(1)
gssfit <- rpart(confed ~ sex + age + race + relig + region + income + educ + incumbency + pres + polviews, data=gss_train, method="class")
fancyRpartPlot(gssfit, main = "Modern Era Classification Tree for Confed")

# Testing and Validation
pred <- predict(gssfit, gss_test, type = "class")
mat <- confusionMatrix(gss_test$confed, pred)
mat
```

As you can see, the classification tree method does not find any viable splits when only fitting on the prior era data. However, when fitting on the modern era data, the method fits a model with an accuracy of `r mat$overall[['Accuracy']]`. Therefore, we can see that it is easier to predict how confident a person is in the executice branch today than in the 70's and 80's.

# Confidence in the Federal Government Logistic Regression

## Confed with Age, Income, and Education

```{r, echo=TRUE}
gss$polviews <- as.integer(gss$polviews) - 1
gss$confed <- as.integer(gss$confed) - 1
gss$pres <- as.integer(gss$pres) - 1



model_a_i_e <- glm(confed ~ age + income + educ, data = gss, family = binomial)
summary(model_a_i_e)

1 - (model_a_i_e$deviance / model_a_i_e$null.deviance)

gss$predicted_prob_a_i_e <- predict(model_a_i_e, type = "response")
gss$predicted_class_a_i_e <- ifelse(gss$predicted_prob_a_i_e > 0.5, 1, 0)

roc_a_i_e <- roc(confed ~ predicted_prob_a_i_e, data = gss)
roc_a_i_e$auc
plot(roc_a_i_e)

gss$predicted_class_a_i_e <- factor(gss$predicted_class_a_i_e, levels = c(0, 1))
gss$confed <- factor(gss$confed, levels = c(0, 1))

conf_matrix_a_i_e <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i_e),
  reference = as.factor(gss$confed)
)
conf_matrix_a_i_e

```

## Confed with Age, Income, Education, and Who R Voted For

```{r, echo=TRUE}
gss$confed <- factor(gss$confed, levels = c(0, 1))
gss$pres <- factor(gss$pres, levels = c(0, 1))

gss <- na.omit(gss)

model_a_i_e_p <- glm(confed ~ age + income + educ + pres, data = gss, family = binomial)
summary(model_a_i_e_p)


1 - (model_a_i_e_p$deviance / model_a_i_e_p$null.deviance)

gss$predicted_prob_a_i_e_p <- predict(model_a_i_e_p, type = "response")
gss$predicted_class_a_i_e_p <- ifelse(gss$predicted_prob_a_i_e_p > 0.5, 1, 0)

roc_a_i_e_p <- roc(confed ~ predicted_prob_a_i_e_p, data = gss)
roc_a_i_e_p$auc
plot(roc_a_i_e_p)

gss$predicted_class_a_i_e_p <- factor(gss$predicted_class_a_i_e_p, levels = c(0, 1))
conf_matrix_a_i_e_p <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i_e_p),
  reference = as.factor(gss$confed)
)
conf_matrix_a_i_e_p

```

```{r, echo=TRUE}
gss_clean <- gss[complete.cases(gss[, c("confed", "age", "income", "educ", "pres")]), ]

model_a_i_e <- glm(confed ~ age + income + educ, data = gss_clean, family = binomial)
model_a_i_e_p <- glm(confed ~ age + income + educ + pres, data = gss_clean, family = binomial)

anova(model_a_i_e, model_a_i_e_p, test = "Chisq")

AIC(model_a_i_e, model_a_i_e_p)

BIC(model_a_i_e, model_a_i_e_p)

logLik_model_a_i_e <- logLik(model_a_i_e)
logLik_model_null_a_i_e <- logLik(glm(confed ~ 1, data = gss_clean, family = binomial))
mcfadden_r2_a_i_e <- 1 - (as.numeric(logLik_model_a_i_e) / as.numeric(logLik_model_null_a_i_e))

logLik_model_a_i_e_p <- logLik(model_a_i_e_p)
logLik_model_null_a_i_e_p <- logLik(glm(confed ~ 1, data = gss_clean, family = binomial))
mcfadden_r2_a_i_e_p <- 1 - (as.numeric(logLik_model_a_i_e_p) / as.numeric(logLik_model_null_a_i_e_p))

mcfadden_r2_a_i_e
mcfadden_r2_a_i_e_p

gss_clean$predicted_prob_a_i_e <- predict(model_a_i_e, type = "response")
gss_clean$predicted_prob_a_i_e_p <- predict(model_a_i_e_p, type = "response")

roc_a_i_e <- roc(confed ~ predicted_prob_a_i_e, data = gss_clean)
roc_a_i_e_p <- roc(confed ~ predicted_prob_a_i_e_p, data = gss_clean)

roc_a_i_e$auc
roc_a_i_e_p$auc

gss_clean$predicted_class_a_i_e <- ifelse(gss_clean$predicted_prob_a_i_e > 0.5, 1, 0)
gss_clean$predicted_class_a_i_e_p <- ifelse(gss_clean$predicted_prob_a_i_e_p > 0.5, 1, 0)

conf_matrix_a_i_e <- confusionMatrix(
  data = as.factor(gss_clean$predicted_class_a_i_e),
  reference = as.factor(gss_clean$confed)
)

conf_matrix_a_i_e_p <- confusionMatrix(
  data = as.factor(gss_clean$predicted_class_a_i_e_p),
  reference = as.factor(gss_clean$confed)
)

conf_matrix_a_i_e
conf_matrix_a_i_e_p

```


# Best Prediction Model of Confed

The analysis reveals that Model 2, which includes the pres variable (representing who respondents voted for), is the best predictive model for confidence in the federal government. This model achieves a lower AIC (2932.4 compared to 2971.6 for Model 1) and a higher AUC (0.6112 compared to 0.5728 for Model 1). The inclusion of the pres variable significantly improves the model's performance, demonstrating that political affiliation and voting behavior are stronger predictors of confidence in the government than demographic factors like age, income, and education alone.

The outputs highlight distinct relationships between demographics and confidence. For example, the coefficient for age in both models is negative, suggesting that older individuals tend to have lower confidence in the government. However, this effect is relatively weak (p = 0.038 in Model 2), reflecting generational differences in trust shaped by historical experiences such as wars, policies, and economic changes. Similarly, the coefficient for educ is positive and highly significant (p < 0.001), indicating that higher education levels correlate with greater confidence. Education's role is likely tied to increased policy awareness, which can polarize trust based on alignment with government actions. Income, on the other hand, shows weak predictive power (p = 0.999 in Model 1, p = 0.653 in Model 2), suggesting that economic factors alone do not strongly influence trust in government.

Importantly, the inclusion of the pres variable reveals that political alignment is a critical factor. The coefficient for pres (0.562 in Model 2, p < 0.001) indicates that individuals are significantly more confident in the government when their preferred candidate is in power. This finding is consistent with the higher AUC in Model 2, demonstrating that voter trust is less influenced by demographics and more by political representation. Events such as economic crises or landmark policies likely amplify these effects, as trust becomes polarized along party lines. The models underscore that while demographics like age and education shape general tendencies, ideological alignment is the dominant driver of confidence in governance. This aligns with broader trends of partisan divides influencing trust, voter turnout, and loyalty across time.


1. **Political Preference Drives Confidence:**  
   Voters are more confident when their chosen leaders are in power, indicating a strong relationship between trust in government and political affiliation.

2. **Demographic Factors are Secondary:**  
   While demographics shape general tendencies, they do not independently predict confidence or voting patterns effectively without incorporating ideological alignment.

3. **Generational Shifts:**  
   Older generations may retain confidence based on historical affiliations eg trust in government institutions while younger generations might be influenced by recent political events or media exposure.

4. **Education's Role in Polarization:**  
   Higher education levels might create polarized views where confidence depends heavily on how closely government actions align with voters' ideals.

5. **Economic Context Matters:**  
   Though income did not strongly predict confidence, external events like recessions or tax policies likely influenced voting motivations particularly in periods of economic instability.




# Political Views Trees

Now lets repeat the process with polviwes as our response variable.


```{r}
gss$polviews <- factor(gss$polviews, levels = c(0, 1), labels = c("Cons", "Lib"))
gss$confed <- factor(gss$confed, levels = c(0, 1), labels = c("Trust", "No Trust"))
gss$pres <- factor(gss$pres, levels = c(0, 1), labels = c("Rep", "Dem"))


# Split training and test data sets
gss_split <- initial_split(gss[gss$era == 0,], prop = 0.70, strata = polviews)
gss_test <- testing(gss_split)
gss_train <- training(gss_split)

#Build Tree using Political Categories
set.seed(1)
gssfit <- rpart(polviews ~ sex + age + race + relig + region + income + educ + incumbency + pres + confed, data=gss_train, method="class")
fancyRpartPlot(gssfit, main = "Prior Era Classification Tree for Polviews")

# Testing and Validation
pred <- predict(gssfit, gss_test, type = "class")
pmat0 <- confusionMatrix(gss_test$polviews, pred)
pmat0
```


```{r}
# Split training and test data sets
gss_split <- initial_split(gss[gss$era == 1,], prop = 0.70, strata = confed)
gss_test <- testing(gss_split)
gss_train <- training(gss_split)

#Build Tree using Political Categories
set.seed(1)
gssfit <- rpart(polviews ~ sex + age + race + relig + region + income + educ + incumbency + pres + confed, data=gss_train, method="class")
summary(gssfit)
fancyRpartPlot(gssfit, main = "Modern Era Classification Tree for Polviews")

# Testing and Validation
pred <- predict(gssfit, gss_test, type = "class")
pmat1 <- confusionMatrix(gss_test$polviews, pred)
pmat1
```


Looking at our results, we can see that the tree method fitted a tree for both time periods. However there differences in shape are interesting. For the model fitted on the prior era data, the tree makes splits at several nodes. However, the model fitted on the current era only made one split, that split being on the pres variable indicating that there is a strong correlation between how a person identifies politically and who they voted for. Let's compare the accuracy of each model. The model for the older data has an accuracy of `r pmat0$overall[['Accuracy']]`, while the model for the new data has an accuracy of `r pmat1$overall[['Accuracy']]`. We can see that the newer model has a much higher accuracy while only slpitting on one variable. This is a strong indication that the diversity of political identity among voters has shrunk over the years.


# Polviews Logistic Regression

As seen below, the bivariate logistic regressions reveal varying degrees of predictive power for the relationship between polviews (political views) and the individual predictors era (year), age, education, and income. The analysis of era shows that it is statistically significant (p < 0.001) but has low predictive power, as indicated by a pseudo-R² of 0.0036 and an AUC of 0.5352. Despite its limitations, era captures temporal effects and should be included in the multivariate model to account for shifts in political views over time. Similarly, age has a significant negative relationship with political views (p < 0.001), suggesting that older individuals are less likely to hold the coded political view. With a pseudo-R² of 0.0155 and an AUC of 0.5845, age demonstrates modest predictive power and is an important demographic variable to include.

Education emerges as one of the strongest predictors in the bivariate analysis, showing a significant positive association with political views (p < 0.001). Individuals with higher levels of education are more likely to hold the coded political view, with a pseudo-R² of 0.0138 and an AUC of 0.5929, slightly outperforming age. Including education in the multivariate model is essential due to its strong relationship with political identity and its potential interactions with other socioeconomic factors. In contrast, income does not significantly predict political views (p = 0.225), with a negligible pseudo-R² of 0.0002 and an AUC of 0.5093. This suggests that income is not a key independent predictor and can be excluded from the multivariate model unless theoretical or interaction-based justifications arise.

In summary, the best variables to include in a multivariate prediction model for political views are era, age, and education. These predictors capture temporal, demographic, and socioeconomic dimensions that are essential for understanding shifts in political views. Income, on the other hand, demonstrates weak predictive power and is unlikely to add value to the model. When we combine the strongest predictors, the multivariate model can provide a comprehensive framework for exploring the factors that shape political views over time.

## Polviews and Year

```{r, echo=TRUE}
gss$polviews <- as.integer(gss$polviews) - 1
gss$confed <- as.integer(gss$confed) - 1
gss$pres <- as.integer(gss$pres) - 1


model_era <- glm(polviews ~ era, data = gss, family = binomial)
summary(model_era)


1 - (model_era$deviance / model_era$null.deviance)

gss$predicted_prob_era <- predict(model_era, type = "response")
gss$predicted_class_era <- ifelse(gss$predicted_prob_era > 0.5, 1, 0)

roc_era <- roc(polviews ~ predicted_prob_era, data = gss)
roc_era$auc
plot(roc_era)

conf_matrix_era <- confusionMatrix(
  data = as.factor(gss$predicted_class_era),
  reference = as.factor(gss$polviews)
)
conf_matrix_era

```

## Polviews and Age

```{r, echo=TRUE}

model_age <- glm(polviews ~ age, data = gss, family = binomial)
summary(model_age)


1 - (model_age$deviance / model_age$null.deviance)

gss$predicted_prob_age <- predict(model_age, type = "response")
gss$predicted_class_age <- ifelse(gss$predicted_prob_age > 0.5, 1, 0)

roc_age <- roc(polviews ~ predicted_prob_age, data = gss)
roc_age$auc
plot(roc_age)

conf_matrix_age <- confusionMatrix(
  data = as.factor(gss$predicted_class_age),
  reference = as.factor(gss$polviews)
)
conf_matrix_age

```

## Polviews and Education

```{r, echo=TRUE}

model_educ <- glm(polviews ~ educ, data = gss, family = binomial)
summary(model_educ)


1 - (model_educ$deviance / model_educ$null.deviance)

gss$predicted_prob_educ <- predict(model_educ, type = "response")
gss$predicted_class_educ <- ifelse(gss$predicted_prob_educ > 0.5, 1, 0)

roc_educ <- roc(polviews ~ predicted_prob_educ, data = gss)
roc_educ$auc
plot(roc_educ)

conf_matrix_educ <- confusionMatrix(
  data = as.factor(gss$predicted_class_educ),
  reference = as.factor(gss$polviews)
)
conf_matrix_educ

```

## Polviews and Income

```{r, echo=TRUE}

model_income <- glm(polviews ~ income, data = gss, family = binomial)
summary(model_income)


1 - (model_income$deviance / model_income$null.deviance)

gss$predicted_prob_income <- predict(model_income, type = "response")
gss$predicted_class_income <- ifelse(gss$predicted_prob_income > 0.5, 1, 0)

roc_income <- roc(polviews ~ predicted_prob_income, data = gss)
roc_income$auc
plot(roc_income)

conf_matrix_income <- confusionMatrix(
  data = as.factor(gss$predicted_class_income),
  reference = as.factor(gss$polviews)
)
conf_matrix_income
```

## Polviews with Age and Income

```{r, echo=TRUE}

model_a_i <- glm(polviews ~ age + income, data = gss, family = binomial)
summary(model_a_i)

1 - (model_a_i$deviance / model_a_i$null.deviance)

gss$predicted_prob_a_i <- predict(model_a_i, type = "response")
gss$predicted_class_a_i <- ifelse(gss$predicted_prob_a_i > 0.5, 1, 0)

roc_a_i <- roc(polviews ~ predicted_prob_a_i, data = gss)
roc_a_i$auc
plot(roc_a_i)

conf_matrix_a_i <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_i
```

# Best Prediction Model for Polviews

This logistic regression model examines the relationship between polviews (political views, with 0 for Republicans and 1 for Democrats) and three predictors: age, era, and educ (education). All three predictors are statistically significant, with p-values below 0.001, indicating their strong association with political affiliation. The pseudo-R² of 0.022 suggests modest explanatory power, and the Area Under the Curve (AUC) of 0.6243 indicates a fair ability to predict whether someone identifies as a Democrat or Republican.

The coefficients provide valuable insights into the predictors' relationships with political affiliation. The intercept is -0.6155, representing the log-odds of identifying as a Democrat for someone in the earlier year span (era = 0), at the average age and education level. The coefficient for age is -0.0200, meaning that for every additional year of age, the log-odds of identifying as a Democrat decrease by 0.02. This suggests that older individuals are slightly more likely to identify as Republican, consistent with generational differences in political leanings. The coefficient for era is 0.2544, indicating that individuals surveyed in the more modern year span are more likely to identify as Democrats, reflecting societal shifts such as increasing political polarization and demographic changes. Finally, the coefficient for educ is 0.0885, suggesting that higher education levels increase the likelihood of identifying as a Democrat, consistent with established trends linking education to liberal political ideologies.

While the model captures meaningful relationships and provides valuable insights, its predictive performance leaves room for improvement. The pseudo-R² indicates that only a small proportion of the variation in political views is explained by the predictors, and the AUC of 0.6243 suggests limited discriminatory power. The accuracy of 60.44%, as shown in the confusion matrix, is only modestly better than random guessing, indicating that the model could benefit from additional predictors or interactions.

To improve the model, additional predictors such as race, geographic region, or household income could be included, as these factors are often strongly associated with political views. Interaction terms, such as between era and educ or age and era, could also capture more nuanced relationships. For instance, the impact of education on political views may differ between the earlier and modern time spans. While the model provides a solid foundation and highlights important trends, expanding and refining the set of predictors could significantly enhance its explanatory and predictive power.

## Logistic Regression of Polviews with Age, Year, and Education

```{r, echo=TRUE}

model_a_e_e <- glm(polviews ~ age + era + educ, data = gss, family = binomial)
summary(model_a_e_e)

1 - (model_a_e_e$deviance / model_a_e_e$null.deviance)

gss$predicted_prob_a_e_e <- predict(model_a_e_e, type = "response")
gss$predicted_class_a_e_e <- ifelse(gss$predicted_prob_a_e_e > 0.5, 1, 0)

roc_a_e_e <- roc(polviews ~ predicted_prob_a_e_e, data = gss)
roc_a_e_e$auc
plot(roc_a_e_e)

conf_matrix_a_e_e <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_e_e),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_e_e

```
