---
title: "Group Project"
author: "Marissa Mowers"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r, echo=TRUE}

library(kableExtra)
library(gssr)
library(gssrdoc)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(ezids)


gss74 <- gss_get_yr(1974)
gss75 <- gss_get_yr(1975)
gss76 <- gss_get_yr(1976)
gss77 <- gss_get_yr(1977)
gss78 <- gss_get_yr(1978)
gss80 <- gss_get_yr(1980)

gss16 <- gss_get_yr(2016)
gss18 <- gss_get_yr(2018)
gss21 <- gss_get_yr(2021)
gss22 <- gss_get_yr(2022)
```

```{r, echo=TRUE}

common_74_80 <- Reduce(intersect, list(colnames(gss74), colnames(gss75), colnames(gss76), colnames(gss77), colnames(gss78), colnames(gss80)))

common_16_22 <- Reduce(intersect, list(colnames(gss16), colnames(gss18), colnames(gss21), colnames(gss22)))                       

gss74_80 <- rbind(gss74 [common_74_80], gss75 [common_74_80], gss76 [common_74_80], gss77 [common_74_80], gss78 [common_74_80], gss80 [common_74_80])

gss16_22 <- rbind(gss16 [common_16_22], gss18 [common_16_22], gss21 [common_16_22], gss22 [common_16_22])

#New datasets
gss74_80

gss16_22 

```


# EDA of Confidence in the Executive Branch of the Federal Government

## EDA of Confidence in the Executive Branch of the Federal Government from 1974 through 1980.

```{r, echo=TRUE}
?confed

xkablesummary(gss74_80[, c("confed")])
missing_cases_confed_74_80 <- sum(is.na(gss74_80$confed))
valid_cases_confed_74_80 <- sum(!is.na(gss74_80$confed))

names(which.max(table(gss74_80$confed)))

missing_cases_confed_74_80
valid_cases_confed_74_80


gss74_80$confed <- factor(gss74_80$confed, levels = c(1, 2, 3), labels = c("A Great Deal", "Only Some", "Hardly Any"))

str(gss16_22$confed)

ggplot(data=gss74_80, aes(x = confed, color=confed, fill=confed)) + 
  geom_bar() +
  labs(x= "Confidence in the Executive Branch of the Federal Government", y= "Frequency") +
  labs(title = "Bar Graph of Confidence in the Executive Branch from 1974-1980") 

```

## EDA of Confidence in the Executive Branch of the Federal Government from 2016 through 2022.
```{r, echo=TRUE}
xkablesummary(gss16_22[, c("confed")])
missing_cases_confed_16_22 <- sum(is.na(gss16_22$confed))
valid_cases_confed_16_22 <- sum(!is.na(gss16_22$confed))
names(which.max(table(gss16_22$confed)))


missing_cases_confed_16_22
valid_cases_confed_16_22


gss16_22$confed <- factor(gss16_22$confed, levels = c(1, 2, 3), labels = c("A Great Deal", "Only Some", "Hardly Any"))

ggplot(data=gss16_22, aes(x = confed, color=confed, fill=confed)) + 
  geom_bar() +
  labs(x= "Confidence in the Executive Branch of the Federal Government", y= "Frequency") +
  labs(title = "Bar Graph of Confidence in the Executive Branch from 2016-2022") 
```


# EDA of Sex

## EDA of Sex from 1974 through 1980.


```{r, echo=TRUE}
xkablesummary(gss74_80[, c("sex")])
missing_cases_sex_74_80 <- sum(is.na(gss74_80$sex))
valid_cases_sex_74_80 <- sum(!is.na(gss74_80$sex))
names(which.max(table(gss74_80$sex)))


missing_cases_sex_74_80
valid_cases_sex_74_80

gss74_80$sex <- factor(gss74_80$sex, levels = c(1, 2), labels = c("Male", "Female"))

str(gss74_80$sex)

ggplot(data=gss74_80, aes(x = sex, color=sex, fill=sex)) + 
  geom_bar() +
  labs(x= "Sex", y= "Frequency") +
  labs(title = "Bar Graph of Sex Distribution from 1974 Through 1980") 
```

## EDA of Sex from 2016 through 2022.


```{r, echo=TRUE}
xkablesummary(gss16_22[, c("sex")])
missing_cases_sex_16_22 <- sum(is.na(gss16_22$sex))
valid_cases_sex_16_22 <- sum(!is.na(gss16_22$sex))
missing_cases_sex_16_22
valid_cases_sex_16_22
names(which.max(table(gss16_22$sex)))


gss16_22$sex <- factor(gss16_22$sex, levels = c(1, 2), labels = c("Male", "Female"))

str(gss16_22$sex)

ggplot(data=gss16_22, aes(x = sex, color=sex, fill=sex)) + 
  geom_bar() +
  labs(x= "Sex", y= "Frequency") +
  labs(title = "Bar Graph of Sex Distribution from 2016 Through 2022") 
```


# EDA of Race

## EDA of Race from 1974 through 1980.


```{r, echo=TRUE}
xkablesummary(gss74_80[, c("race")])
missing_cases_race_74_80 <- sum(is.na(gss74_80$race))
valid_cases_race_74_80 <- sum(!is.na(gss74_80$race))
names(which.max(table(gss74_80$race)))


missing_cases_race_74_80
valid_cases_race_74_80

gss74_80$race <- factor(gss74_80$race, levels = c(1, 2, 3), labels = c("White", "Black", "Other"))

str(gss74_80$race)

ggplot(data=gss74_80, aes(x = race, color=race, fill=race)) + 
  geom_bar() +
  labs(x= "Race", y= "Frequency") +
  labs(title = "Bar Graph of Race Distribution from 1974 Through 1980") 
```


## EDA of Race from 2016 through 2022.


```{r, echo=TRUE}
xkablesummary(gss16_22[, c("race")])
missing_cases_race_16_22 <- sum(is.na(gss16_22$race))
valid_cases_race_16_22 <- sum(!is.na(gss16_22$race))
names(which.max(table(gss16_22$race)))


missing_cases_race_16_22
valid_cases_race_16_22

gss16_22$race <- factor(gss16_22$race, levels = c(1, 2, 3), labels = c("White", "Black", "Other"))

str(gss16_22$race)

ggplot(data=gss16_22, aes(x = race, color=race, fill=race)) + 
  geom_bar() +
  labs(x= "Race", y= "Frequency") +
  labs(title = "Bar Graph of Race Distribution from 2016 Through 2022") 
```

# Chi-Squared Test 



#Confidence in the Executive Branch of the Federal Government vs. Sex

## 1974 through 1980


```{r, echo=TRUE}
confed_sex_table <- table(gss74_80$confed, gss74_80$sex)

confed_sex_chi <- chisq.test(confed_sex_table)

kable(confed_sex_table, caption = "Contingency Table: Confidence in Executive Branch vs. Sex (1974-1980)") %>%
  kable_styling(full_width = FALSE)

confed_sex_chi

```


## 2016 through 2022


```{r, echo=TRUE}
confed_sex_table <- table(gss16_22$confed, gss16_22$sex)

confed_sex_chi <- chisq.test(confed_sex_table)

kable(confed_sex_table, caption = "Contingency Table: Confidence in Executive Branch vs. Sex") %>%
  kable_styling(full_width = FALSE)

confed_sex_chi

```

# Confidence in the Executive Branch of the Federal Government vs. Race

## 1974 through 1980


```{r, echo=TRUE}
confed_race_table <- table(gss74_80$confed, gss74_80$race)

confed_race_chi <- chisq.test(confed_race_table)

kable(confed_race_table, caption = "Contingency Table: Confidence in Executive Branch vs. Race (1974-1980)") %>%
  kable_styling(full_width = FALSE)

confed_race_chi

```

## 2016 through 2022


```{r, echo=TRUE}
confed_race_table <- table(gss16_22$confed, gss16_22$race)
confed_race_chi <- chisq.test(confed_race_table)

kable(confed_race_table, caption = "Contingency Table: Confidence in Executive Branch vs. Race") %>%
  kable_styling(full_width = FALSE)

confed_race_chi

```


# Confidence in the Executive Branch of the Federal Government vs. Religion

## 1974 through 1980


```{r, echo=TRUE}
confed_religion_table <- table(gss74_80$confed, gss74_80$relig)

confed_religion_chi <- chisq.test(confed_religion_table)

kable(confed_religion_table, caption = "Contingency Table: Confidence in Executive Branch vs. Religion (1974-1980)") %>%
  kable_styling(full_width = FALSE)

confed_religion_chi

```

## 2016 through 2022


```{r, echo=TRUE}
confed_religion_table <- table(gss16_22$confed, gss16_22$relig)
confed_religion_chi <- chisq.test(confed_religion_table)

kable(confed_religion_table, caption = "Contingency Table: Confidence in Executive Branch vs. Religion") %>%
  kable_styling(full_width = FALSE)

confed_religion_chi

```

```{r, echo=TRUE}
gss74_80$relig <- factor(gss74_80$relig, 
                            levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13), 
                            labels = c("Protestant", "Catholic", "Jewish", "None", "Other", "Other", "Other", "Other", "Other", "Other", "Other", "Other", "Other"))

table(gss74_80$relig)

gss16_22$relig <- factor(gss16_22$relig, 
                            levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13), 
                            labels = c("Protestant", "Catholic", "Jewish", "None", "Other", "Other", "Other", "Other", "Other", "Other", "Other", "Other", "Other"))

table(gss16_22$relig)
```


# Confidence in the Executive Branch of the Federal Government vs. Religion - RECODED

## 1974 through 1980 - RECODED


```{r, echo=TRUE}
confed_religion_table <- table(gss74_80$confed, gss74_80$relig)

confed_religion_chi <- chisq.test(confed_religion_table)

kable(confed_religion_table, caption = "Contingency Table: Confidence in Executive Branch vs. Religion (1974-1980)") %>%
  kable_styling(full_width = FALSE)

confed_religion_chi

```


## 2016 through 2022 - RECODED


```{r, echo=TRUE}
confed_religion_table <- table(gss16_22$confed, gss16_22$relig)
confed_religion_chi <- chisq.test(confed_religion_table)

kable(confed_religion_table, caption = "Contingency Table: Confidence in Executive Branch vs. Religion") %>%
  kable_styling(full_width = FALSE)

confed_religion_chi

```

#EDA for Vaijayanti
```{r, echo=TRUE}
sum(is.na(gss74_80$age))
sum(!is.na(gss74_80$age))
names(which.max(table(gss74_80$age)))
xkablesummary(gss74_80[, c("age")])


sum(is.na(gss16_22$age))
sum(!is.na(gss16_22$age))
names(which.max(table(gss16_22$age)))
xkablesummary(gss16_22[, c("age")])


sum(is.na(gss74_80$income))
sum(!is.na(gss74_80$income))
names(which.max(table(gss74_80$income)))
xkablesummary(gss74_80[, c("income")])


sum(is.na(gss16_22$income16))
sum(!is.na(gss16_22$income16))
names(which.max(table(gss16_22$income16)))
xkablesummary(gss16_22[, c("income16")])

 

```
