---
title: "Group Project #2"
author: " Joshua Gray, Vaijayanti Deshmukh, & Marissa Mowers"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: 3
---


```{r, echo=FALSE, result='hide'}
library(kableExtra)
library(gssr)
library(gssrdoc)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(ezids)
library(RColorBrewer)
library(psych)
library(broom)
library(dplyr)
library(pROC)
library(caret)
library(rattle)
library(rsample)

file_path <- "C:\\Users\\mnmow\\OneDrive\\Desktop\\GWU\\Fall 2024\\DATS6101 Intro\\DATS-6101-Project\\Project 2\\GSS_data.csv"
gss <- read.csv(file_path)

str(gss)

gss <- na.omit(gss)

```

# Polviews bivariate graph

```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(polviews))) +
  geom_bar(position = "dodge") + 
  scale_fill_manual(
    values = c("0" = "red", "1" = "blue"), 
    labels = c("Republican", "Democrat"), 
    name = "Political View"
  ) +
  labs(
    title = "Frequency of Political Views Over the Years",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Polviews bivariate graph

```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(confed))) +
  geom_bar(position = "dodge") + 
  scale_fill_manual(
    values = c("0" = "lightgreen", "1" = "orchid"), 
    labels = c("Trust", "No Trust"), 
    name = "Confidence in the Executive Branch"
  ) +
  labs(
    title = "Frequency of Confidence in the Executive Branch Over the Years",
    x = "Years",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Tetrachoric Correlation 

```{r, echo=FALSE}


binary_vars <- gss %>% select(polviews, pres, incumbency)

binary_vars <- binary_vars %>% mutate_all(as.numeric) %>% na.omit()
lapply(binary_vars, unique)

tetra_corr <- psych::tetrachoric(binary_vars)
print(tetra_corr$rho)

```

# Logistic Regression

## Polviews and Year

```{r, echo=TRUE}

model_era <- glm(polviews ~ era, data = gss, family = binomial)
summary(model_era)


1 - (model_era$deviance / model_era$null.deviance)

gss$predicted_prob_era <- predict(model_era, type = "response")
gss$predicted_class_era <- ifelse(gss$predicted_prob_era > 0.5, 1, 0)

roc_era <- roc(polviews ~ predicted_prob_era, data = gss)
roc_era$auc
plot(roc_era)

conf_matrix_era <- confusionMatrix(
  data = as.factor(gss$predicted_class_era),
  reference = as.factor(gss$polviews)
)
conf_matrix_era

```

## Polviews and Age

```{r, echo=TRUE}

model_age <- glm(polviews ~ age, data = gss, family = binomial)
summary(model_age)


1 - (model_age$deviance / model_age$null.deviance)

gss$predicted_prob_age <- predict(model_age, type = "response")
gss$predicted_class_age <- ifelse(gss$predicted_prob_age > 0.5, 1, 0)

roc_age <- roc(polviews ~ predicted_prob_age, data = gss)
roc_age$auc
plot(roc_age)

conf_matrix_age <- confusionMatrix(
  data = as.factor(gss$predicted_class_age),
  reference = as.factor(gss$polviews)
)
conf_matrix_age

```

## Polviews and Education

```{r, echo=TRUE}

model_educ <- glm(polviews ~ educ, data = gss, family = binomial)
summary(model_educ)


1 - (model_educ$deviance / model_educ$null.deviance)

gss$predicted_prob_educ <- predict(model_educ, type = "response")
gss$predicted_class_educ <- ifelse(gss$predicted_prob_educ > 0.5, 1, 0)

roc_educ <- roc(polviews ~ predicted_prob_educ, data = gss)
roc_educ$auc
plot(roc_educ)

conf_matrix_educ <- confusionMatrix(
  data = as.factor(gss$predicted_class_educ),
  reference = as.factor(gss$polviews)
)
conf_matrix_educ

```

## Polviews and Income

```{r, echo=TRUE}

model_income <- glm(polviews ~ income, data = gss, family = binomial)
summary(model_income)


1 - (model_income$deviance / model_income$null.deviance)

gss$predicted_prob_income <- predict(model_income, type = "response")
gss$predicted_class_income <- ifelse(gss$predicted_prob_income > 0.5, 1, 0)

roc_income <- roc(polviews ~ predicted_prob_income, data = gss)
roc_income$auc
plot(roc_income)

conf_matrix_income <- confusionMatrix(
  data = as.factor(gss$predicted_class_income),
  reference = as.factor(gss$polviews)
)
conf_matrix_income
```

## Polviews with Age and Income

```{r, echo=TRUE}

model_a_i <- glm(polviews ~ age + income, data = gss, family = binomial)
summary(model_a_i)

1 - (model_a_i$deviance / model_a_i$null.deviance)

gss$predicted_prob_a_i <- predict(model_a_i, type = "response")
gss$predicted_class_a_i <- ifelse(gss$predicted_prob_a_i > 0.5, 1, 0)

roc_a_i <- roc(polviews ~ predicted_prob_a_i, data = gss)
roc_a_i$auc
plot(roc_a_i)

conf_matrix_a_i <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_i
```

## Polviews with Age, Income, and Education

```{r, echo=TRUE}

model_a_i_e <- glm(polviews ~ age + income + educ, data = gss, family = binomial)
summary(model_a_i_e)

1 - (model_a_i_e$deviance / model_a_i_e$null.deviance)

gss$predicted_prob_a_i_e <- predict(model_a_i_e, type = "response")
gss$predicted_class_a_i_e <- ifelse(gss$predicted_prob_a_i_e > 0.5, 1, 0)

roc_a_i_e <- roc(polviews ~ predicted_prob_a_i_e, data = gss)
roc_a_i_e$auc
plot(roc_a_i_e)

conf_matrix_a_i_e <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i_e),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_i_e

```

## Polviews with Age, Income, Education, and Who R Voted For

```{r, echo=TRUE}

model_a_i_e_p <- glm(polviews ~ age + income + educ + pres, data = gss, family = binomial)
summary(model_a_i_e_p)

1 - (model_a_i_e_p$deviance / model_a_i_e_p$null.deviance)

gss$predicted_prob_a_i_e_p <- predict(model_a_i_e_p, type = "response")
gss$predicted_class_a_i_e_p <- ifelse(gss$predicted_prob_a_i_e_p > 0.5, 1, 0)

roc_a_i_e_p <- roc(polviews ~ predicted_prob_a_i_e_p, data = gss)
roc_a_i_e_p$auc
plot(roc_a_i_e_p)

conf_matrix_a_i_e_p <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i_e_p),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_i_e_p
```

## Polviews with Age, Income, Education, Who R Voted For, and Year

```{r, echo=TRUE}

model_all <- glm(polviews ~ age + income + educ + pres + era, data = gss, family = binomial)
summary(model_all)


1 - (model_all$deviance / model_all$null.deviance)

gss$predicted_prob_all <- predict(model_all, type = "response")
gss$predicted_class_all <- ifelse(gss$predicted_prob_all > 0.5, 1, 0)

roc_all <- roc(polviews ~ predicted_prob_all, data = gss)
roc_all$auc
plot(roc_all)

conf_matrix_all <- confusionMatrix(
  data = as.factor(gss$predicted_class_all),
  reference = as.factor(gss$polviews)
)
conf_matrix_all
```

## Polviews with Age, Income, Education, and an Interaction Between Who R Voted For and Incumbency

```{r, echo=TRUE}

interaction_model_wo_era <- glm(polviews ~ age + income + educ + pres * incumbency, data = gss, family = binomial)
summary(interaction_model_wo_era)

1 - (interaction_model_wo_era$deviance / interaction_model_wo_era$null.deviance)

gss$predicted_prob_wo_era <- predict(interaction_model_wo_era, type = "response")
gss$predicted_class_wo_era <- ifelse(gss$predicted_prob_wo_era > 0.5, 1, 0)

roc_wo_era <- roc(polviews ~ predicted_prob_wo_era, data = gss)
roc_wo_era$auc
plot(roc_wo_era)

conf_matrix_wo_era <- confusionMatrix(
  data = as.factor(gss$predicted_class_wo_era),
  reference = as.factor(gss$polviews)
)
conf_matrix_wo_era
```

## Polviews with Age, Income, Education, Year, and an Interaction Between Who R Voted For and Incumbency

```{r, echo=TRUE}

interaction_model_era <- glm(polviews ~ age + income + educ + era + pres * incumbency, data = gss, family = binomial)
summary(interaction_model_era)


1 - (interaction_model_era$deviance / interaction_model_era$null.deviance)

gss$predicted_prob_era_interaction <- predict(interaction_model_era, type = "response")
gss$predicted_class_era_interaction <- ifelse(gss$predicted_prob_era_interaction > 0.5, 1, 0)

roc_era_interaction <- roc(polviews ~ predicted_prob_era_interaction, data = gss)
roc_era_interaction$auc
plot(roc_era_interaction)

conf_matrix_era_interaction <- confusionMatrix(
  data = as.factor(gss$predicted_class_era_interaction),
  reference = as.factor(gss$polviews)
)
conf_matrix_era_interaction
```




```{r, echo=TRUE}
str(gss)

# Levels of measurement:
# polviews: Nominal (categorical variable with two possible values)
# age: Ratio (continuous variable)
# income: Ratio (continuous variable)
# educ: Ordinal (categorical, education levels)
# era: Ordinal (categorical, time period)
# pres: Nominal (categorical, presence of president)
# incumbency: Nominal (categorical, whether the person is an incumbent)

# Load necessary libraries
library(knitr)
library(dplyr)
library(ezids)

xkablesummary(data.frame(confed = gss$confed))
xkablesummary(data.frame(polviews = gss$polviews))
xkablesummary(data.frame(age = gss$age))
xkablesummary(data.frame(income = gss$income))
xkablesummary(data.frame(educ = gss$educ))
xkablesummary(data.frame(era = gss$era))
xkablesummary(data.frame(pres = gss$pres))
xkablesummary(data.frame(incumbency = gss$incumbency))
```

