

```markdown
---
title: "Data Science Project 1"
author: "Vaijayanti Deshmukh"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r, echo=FALSE}
library(kableExtra)
library(gssr)
library(ggplot2)
library(RColorBrewer)
library(knitr)

# Load individual years
gss74 <- gss_get_yr(1974)
gss75 <- gss_get_yr(1975)
gss76 <- gss_get_yr(1976)
gss77 <- gss_get_yr(1977)
gss78 <- gss_get_yr(1978)
gss80 <- gss_get_yr(1980)

gss16 <- gss_get_yr(2016)
gss18 <- gss_get_yr(2018)
gss21 <- gss_get_yr(2021)
gss22 <- gss_get_yr(2022)
```

```{r, echo=FALSE}
# Find common columns for 1974-1980 datasets
common_74_80 <- Reduce(intersect, list(colnames(gss74), colnames(gss75), colnames(gss76), colnames(gss77), colnames(gss78), colnames(gss80)))

# Find common columns for 2016-2022 datasets
common_16_22 <- Reduce(intersect, list(colnames(gss16), colnames(gss18), colnames(gss21), colnames(gss22)))

# Subset the datasets to keep only common columns
gss74_80 <- do.call(rbind, lapply(list(gss74, gss75, gss76, gss77, gss78, gss80), function(x) x[, common_74_80]))
gss16_22 <- do.call(rbind, lapply(list(gss16, gss18, gss21, gss22), function(x) x[, common_16_22]))

# Display new datasets
gss74_80
gss16_22
```

```{r ,echo=FALSE}
# Check column names for each dataset
colnames(gss16)
colnames(gss18)
colnames(gss21)
colnames(gss22)

# Check the number of columns in each dataset
sapply(list(gss16, gss18, gss21, gss22), ncol)

# Find common columns for the 2016-2022 datasets
common_cols_16_22 <- Reduce(intersect, lapply(list(gss16, gss18, gss21, gss22), colnames))

# Function to select common columns
select_common_columns <- function(df, common_cols) {
  return(df[, common_cols, drop = FALSE])
}

# Combine datasets for 2016-2022
gss16_22 <- do.call(rbind, lapply(list(gss16, gss18, gss21, gss22), function(x) select_common_columns(x, common_cols_16_22)))

# Check the structure of the combined dataset
str(gss16_22)

# Check for NA values
sum(is.na(gss16_22))
```

# Age Analysis

```{r AGE, echo=TRUE}
# Checking summary of the age variable for 1974-1980
summary(gss74_80$age)

# Visualizing age distribution using ggplot2 for 1974-1980
ggplot(gss74_80, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(title = "Age Distribution (1974-1980)", x = "Age", y = "Frequency") +
  theme_minimal()

# Checking summary of the age variable for 2016-2022
summary(gss16_22$age)

# Visualizing age distribution using ggplot2 for 2016-2022
ggplot(gss16_22, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "lightgreen", color = "black") +
  labs(title = "Age Distribution (2016-2022)", x = "Age", y = "Frequency") +
  theme_minimal()

# Creating age groups for analysis
gss74_80$age_group <- cut(gss74_80$age, breaks = c(0, 25, 45, 65, 100), 
                          labels = c("Under 25", "25-45", "45-65", "65+"), right = FALSE)
gss16_22$age_group <- cut(gss16_22$age, breaks = c(0, 25, 45, 65, 100), 
                          labels = c("Under 25", "25-45", "45-65", "65+"), right = FALSE)

# Conducting chi-squared test for age groups
confed_age_table <- table(gss74_80$confed, gss74_80$age_group)
confed_age_chi <- chisq.test(confed_age_table)

# Displaying the contingency table and chi-squared results
kable(confed_age_table, caption = "Confed vs. Age Group (1974-1980)") %>%
  kable_styling(full_width = FALSE)

confed_age_chi
```

# Education Analysis

```{r EDU, echo=TRUE}
# Checking summary statistics for education for 1974-1980
summary(gss74_80$educ)

# Plotting education levels for 1974-1980
ggplot(gss74_80, aes(x = educ)) +
  geom_histogram(binwidth = 1, fill = "cornflowerblue", color = "black") +
  labs(title = "Education Distribution (1974-1980)", x = "Years of Education", y = "Frequency") +
  theme_minimal()

# Checking summary statistics for education for 2016-2022
summary(gss16_22$educ)

# Plotting education levels for 2016-2022
ggplot(gss16_22, aes(x = educ)) +
  geom_histogram(binwidth = 1, fill = "palegreen", color = "black") +
  labs(title = "Education Distribution (2016-2022)", x = "Years of Education", y = "Frequency") +
  theme_minimal()

# Creating education groups for analysis
gss74_80$educ_group <- cut(gss74_80$educ, breaks = c(0, 12, 16, 20), 
                           labels = c("High School or Less", "Some College", "Graduate"), right = FALSE)
gss16_22$educ_group <- cut(gss16_22$educ, breaks = c(0, 12, 16, 20), 
                           labels = c("High School or Less", "Some College", "Graduate"), right = FALSE)

# Conducting chi-squared test for education groups
confed_educ_table <- table(gss74_80$confed, gss74_80$educ_group)
confed_educ_chi <- chisq.test(confed_educ_table)

# Displaying the results
kable(confed_educ_table, caption = "Confed vs. Education Group (1974-1980)") %>%
  kable_styling(full_width = FALSE)

confed_educ_chi
```

# Income Analysis
```{r INCOME,echo=TRUE}

# Checking summary statistics for income for 1974-1980
summary(gss74_80$income)

# Visualizing income distribution for 1974-1980
ggplot(gss74_80, aes(x = income)) +
  geom_histogram(binwidth = 1000, fill = "plum", color = "black") +
  labs(title = "Income Distribution (1974-1980)", x = "Income", y = "Frequency") +
  theme_minimal()

# Checking summary statistics for income for 2016-2022
summary(gss16_22$income)

# Visualizing income distribution for 2016-2022
ggplot(gss16_22, aes(x = income)) +
  geom_histogram(binwidth = 1000, fill = "orchid", color = "black") +
  labs(title = "Income Distribution (2016-2022)", x = "Income", y = "Frequency") +
  theme_minimal()

# Grouping income into categories for analysis
gss74_80$income_group <- cut(gss74_80$income, breaks = c(0, 10000, 30000, 50000, 100000, Inf), 
                             labels = c("Low", "Lower-Middle", "Middle", "Upper-Middle", "High"))
gss16_22$income_group <- cut(gss16_22$income, breaks = c(0, 10000, 30000, 50000, 100000, Inf), 
                             labels = c("Low", "Lower-Middle", "Middle", "Upper-Middle", "High"))

# Conducting chi-squared test for income groups
confed_income_table <- table(gss74_80$confed, gss74_80$income_group)
confed_income_chi <- chisq.test(confed_income_table)

print(confed_income_table)


# Displaying the results
kable(confed_income_table, caption = "Confed vs. Income Group (1974-1980)") %>%
  kable_styling(full_width = FALSE)

confed_income_chi
```
