---
title: "EDA for Education"
author: "Vaijayanti Deshmukh"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gssr)
library(gssrdoc)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
```
```{r}
#Read in data frames, filter data
vars <- c('X', 'year', 'id', 'sex', 'age', 'race', 'relig', 'educ', 'income', 'region', 'polviews')
gss_old <- read.csv("C:\\Users\\vaiju\\Documents\\Data Science\\Project DS VJ\\DATS-6101-Project\\gss74_80.csv")
gss_old <- gss_old[, vars]
gss_new <- read.csv("C:\\Users\\vaiju\\Documents\\Data Science\\Project DS VJ\\DATS-6101-Project\\gss16_22.csv")
gss_new <- gss_new[, vars]
```
```{r}
#convert variables to factors with labels
gss_old$region <- factor(gss_old$region, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9), labels = c("new england", "middle atlantic", "east north central", "west north central", "south atlantic", "east south atlantic", "west south central", "mountain", "pacific"))

gss_new$region <- factor(gss_new$region, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9), labels = c("new england", "middle atlantic", "east north central", "west north central", "south atlantic", "east south atlantic", "west south central", "mountain", "pacific"))
```
```{r}
# Change education buckets 
gss_new$educ <- case_when(
  gss_new$educ == 1 ~ 'High School or Less',
  gss_new$educ == 2 ~ 'High School or Less',
  gss_new$educ == 3 ~ 'High School or Less',
  gss_new$educ == 4 ~ 'High School or Less',
  gss_new$educ == 5 ~ 'High School or Less',
  gss_new$educ == 6 ~ 'High School or Less',
  gss_new$educ == 7 ~ 'High School or Less',
  gss_new$educ == 8 ~ 'High School or Less',
  gss_new$educ == 9 ~ 'High School or Less',
  gss_new$educ == 10 ~ 'High School or Less',
  gss_new$educ == 11 ~ 'High School or Less',
  gss_new$educ == 12 ~ 'High School or Less',
  gss_new$educ == 13 ~ 'Some College',
  gss_new$educ == 14 ~ 'Some College',
  gss_new$educ == 15 ~ 'Some College',
  gss_new$educ == 16 ~ 'Some College',
  gss_new$educ == 17 ~ 'Graduate',
  gss_new$educ == 18 ~ 'Graduate',
  gss_new$educ == 19 ~ 'Graduate',
  gss_new$educ == 20 ~ 'Graduate',
  TRUE ~ 'NA'
)


  gss_old$educ <- case_when(
  gss_old$educ == 1 ~ 'High School or Less',
  gss_old$educ == 2 ~ 'High School or Less',
  gss_old$educ == 3 ~ 'High School or Less',
  gss_old$educ == 4 ~ 'High School or Less',
  gss_old$educ == 5 ~ 'High School or Less',
  gss_old$educ == 6 ~ 'High School or Less',
  gss_old$educ == 7 ~ 'High School or Less',
  gss_old$educ == 8 ~ 'High School or Less',
  gss_old$educ == 9 ~ 'High School or Less',
  gss_old$educ == 10 ~ 'High School or Less',
  gss_old$educ == 11 ~ 'High School or Less',
  gss_old$educ == 12 ~ 'High School or Less',
  gss_old$educ == 13 ~ 'Some College',
  gss_old$educ == 14 ~ 'Some College',
  gss_old$educ == 15 ~ 'Some College',
  gss_old$educ == 16 ~ 'Some College',
  gss_old$educ == 17 ~ 'Graduate',
  gss_old$educ == 18 ~ 'Graduate',
  gss_old$educ == 19 ~ 'Graduate',
  gss_old$educ == 20 ~ 'Graduate',
  TRUE ~ 'NA'
)
```
```{r}
# Combine two tables into one
gss_combo <- rbind(gss_old, gss_new)
```

# Education

### Mean, Median, and Mode

The most frequent education in the 70s is `r names(which.max(table(gss_old$educ)))`. 
The most frequent education in the 20s is `r names(which.max(table(gss_new$educ)))`.
The least frequent education in the 70's is `r names(which.min(table(gss_old$educ)))`. 
And the least frequent education in the 20s is `r names(which.min(table(gss_new$educ)))`.

### Distribution

```{r}

library(ggplot2)

# Histogram for gss_old
ggplot(gss_old, aes(x = educ, color = region, fill = region)) + 
  geom_histogram(stat = "Count") + 
  labs(title = "Education 1974-1980", x = "Education Level") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))

# Histogram for gss_new
ggplot(gss_new, aes(x = educ, color = region, fill = region)) + 
  geom_histogram(stat = "Count") + 
  labs(title = "Education 2016-2022", x = "Education Level") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))


# Plot bar charts for education levels in the older dataset (1974-1980)
ggplot(gss_old, aes(x = educ, fill = region)) + 
  geom_bar(color = "blue", position = "dodge") + 
  labs(title = "Education Distribution (1974-1980)", x = "Education Level", y = "Count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10)) +
  scale_fill_brewer(palette = "Set3") 

# Plot bar charts for education levels in the newer dataset (2016-2022)
ggplot(gss_new, aes(x = educ, fill = region)) + 
  geom_bar(color = "blue", position = "dodge") + 
  labs(title = "Education Distribution (2016-2022)", x = "Education Level", y = "Count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10)) +
  scale_fill_brewer(palette = "Set3")  

```


### Mode
The most frequent education in the early surveys is  `r names(which.max(table(gss_old$educ)))`. 
The most frequent education in the recent surveys is `r names(which.max(table(gss_new$educ)))`.



```