---
title: "EDA for Income"
author: "Vaijayanti Deshmukh"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gssr)
library(gssrdoc)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
```
```{r}
#Read in data frames, filter data
vars <- c('X', 'year', 'id', 'sex', 'age', 'race', 'relig', 'educ', 'income', 'region', 'polviews')
gss_old <- read.csv("C:\\Users\\vaiju\\Documents\\Data Science\\Project DS VJ\\DATS-6101-Project\\gss74_80.csv")
gss_old <- gss_old[, vars]
gss_new <- read.csv("C:\\Users\\vaiju\\Documents\\Data Science\\Project DS VJ\\DATS-6101-Project\\gss16_22.csv")
gss_new <- gss_new[, vars]
```


```{r}
# Change income buckets 
gss_new$income <- case_when(
  gss_new$income == 1 ~ 'Low',
  gss_new$income == 2 ~ 'Low',
  gss_new$income== 3 ~ 'Low',
  gss_new$income == 4 ~ 'Lower Middle',
  gss_new$income == 5 ~ 'Lower Middle',
  gss_new$income == 6 ~ 'Middle',
  gss_new$income == 7 ~ 'Middle',
  gss_new$income == 8 ~ 'Middle',
  gss_new$income == 9 ~ 'Upper Middle',
  gss_new$income == 10 ~ 'Upper Middle',
  gss_new$income == 11 ~ 'High',
  gss_new$income == 12 ~ 'High',
  TRUE ~ 'NA'
)


  gss_old$income <- case_when(
  gss_old$income == 1 ~ 'Low',
  gss_old$income == 2 ~ 'Low',
  gss_old$income == 3 ~ 'Low',
  gss_old$income == 4 ~ 'Lower Middle',
  gss_old$income == 5 ~ 'Lower Middle',
  gss_old$income == 6 ~ 'Middle',
  gss_old$income == 7 ~ 'Middle',
  gss_old$income == 8 ~ 'Middle',
  gss_old$income == 9 ~ 'Upper Middle',
  gss_old$income == 10 ~ 'Upper Middle',
  gss_old$income == 11 ~ 'High',
  gss_old$income == 12 ~ 'High',
  TRUE ~ 'NA'
)
```
```{r}
# Combine two tables into one
gss_combo <- rbind(gss_old, gss_new)
```

# Education

### Mean, Median, and Mode

The most frequent Income in the 70s is `r names(which.max(table(gss_old$income)))`. 
The most frequent Income in the 20s is `r names(which.max(table(gss_new$income)))`.
The least frequent Income in the 70's is `r names(which.min(table(gss_old$income)))`. 
And the least frequent Income in the 20s is `r names(which.min(table(gss_new$income)))`.

### Distribution

```{r}

library(ggplot2)

# Histogram for gss_old
ggplot(gss_old, aes(x = income, color = income, fill = income)) + 
  geom_histogram(stat = "Count") + 
  labs(title = "Income 1974-1980", x = "Income Level") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))

# Histogram for gss_new
ggplot(gss_new, aes(x = income, color = income, fill = income)) + 
  geom_histogram(stat = "Count") + 
  labs(title = "Income 2016-2022", x = "Income Level") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))



```


### Mode
The most frequent Income in the early surveys is  `r names(which.max(table(gss_old$income)))`. 
The most frequent Income in the recent surveys is `r names(which.max(table(gss_new$income)))`.



```