---
title: "Project2"
author: "Vaijayanti Deshmukh"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(gssr)
library(gssrdoc)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(ezids)
library(RColorBrewer)
library(psych)
library(broom)
library(dplyr)
library(pROC)
library(caret)

```

```{r}
file_path <- "C:\\Users\\vaiju\\Documents\\Data Science\\Project DS VJ\\Project2 VD\\GSS_data.csv"
gss <- read.csv(file_path)

str(gss)

gss <- na.omit(gss)

```
#EDA for Pres
```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(pres))) +  
  geom_bar(stat = "count", fill = "skyblue") +  
  labs(x = "Pres (Leader Voted)", y = "Frequency", title = "Frequency of Votes for Leader (pres)") +
  theme_minimal()
```
```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(pres))) +
  geom_bar(position = "dodge") +  
  scale_fill_manual(
    values = c("0" = "coral", "1" = "pink"), 
    labels = c("Opponent", "Elected Leader"), 
    name = "Leader Voted"
  ) +
  labs(
    title = "Frequency of Votes for Leader",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#Confed BiVariate Graph

```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(confed))) +
  geom_bar(position = "dodge") + 
  scale_fill_manual(
    values = c("0" = "coral", "1" = "pink"), 
    labels = c("Low", "High"), 
    name = "Confidence in Government"
  ) +
  labs(
    title = "Confidence Over the Years",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
#Confed BiVariate Graph

```{r, echo=TRUE}
ggplot(gss, aes(x = as.factor(year), fill = as.factor(pres))) +
  geom_bar(position = "dodge") + 
  scale_fill_manual(
    values = c("0" = "coral", "1" = "pink"), 
    labels = c("Low", "High"), 
    name = "Confidence in Government"
  ) +
  labs(
    title = "Confidence Over the Years",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#Logistic Regression
##Confed and Year
```{r, echo=TRUE}

model_year <- glm(confed ~ year, data = gss, family = binomial)
summary(model_year)


1 - (model_year$deviance / model_year$null.deviance)

gss$predicted_prob_year <- predict(model_year, type = "response")
gss$predicted_class_year <- ifelse(gss$predicted_prob_year > 0.5, 1, 0)

roc_year <- roc(polviews ~ predicted_prob_year, data = gss)
roc_year$auc
plot(roc_year)

conf_matrix_year <- confusionMatrix(
  data = as.factor(gss$predicted_class_year),
  reference = as.factor(gss$polviews)
)
conf_matrix_year

```

##Confed and Age
```{r, echo=TRUE}

model_age <- glm(confed ~ age, data = gss, family = binomial)
summary(model_age)


1 - (model_age$deviance / model_age$null.deviance)

gss$predicted_prob_age <- predict(model_age, type = "response")
gss$predicted_class_age <- ifelse(gss$predicted_prob_age > 0.5, 1, 0)

roc_age <- roc(polviews ~ predicted_prob_age, data = gss)
roc_age$auc
plot(roc_age)

conf_matrix_age <- confusionMatrix(
  data = as.factor(gss$predicted_class_age),
  reference = as.factor(gss$polviews)
)
conf_matrix_age

```

##Confed and Education
```{r, echo=TRUE}

model_educ <- glm(confed ~ educ, data = gss, family = binomial)
summary(model_educ)


1 - (model_educ$deviance / model_educ$null.deviance)

gss$predicted_prob_educ <- predict(model_educ, type = "response")
gss$predicted_class_educ <- ifelse(gss$predicted_prob_educ > 0.5, 1, 0)

roc_educ <- roc(polviews ~ predicted_prob_educ, data = gss)
roc_educ$auc
plot(roc_educ)

conf_matrix_educ <- confusionMatrix(
  data = as.factor(gss$predicted_class_educ),
  reference = as.factor(gss$polviews)
)
conf_matrix_educ

```

##Confed and Income

```{r, echo=TRUE}

model_income <- glm(confed ~ income, data = gss, family = binomial)
summary(model_income)


1 - (model_income$deviance / model_income$null.deviance)

gss$predicted_prob_income <- predict(model_income, type = "response")
gss$predicted_class_income <- ifelse(gss$predicted_prob_income > 0.5, 1, 0)

roc_income <- roc(polviews ~ predicted_prob_income, data = gss)
roc_income$auc
plot(roc_income)

conf_matrix_income <- confusionMatrix(
  data = as.factor(gss$predicted_class_income),
  reference = as.factor(gss$polviews)
)
conf_matrix_income
```

## Confed with Age and Income

```{r, echo=TRUE}

model_a_i <- glm(confed ~ age + income, data = gss, family = binomial)
summary(model_a_i)

1 - (model_a_i$deviance / model_a_i$null.deviance)

gss$predicted_prob_a_i <- predict(model_a_i, type = "response")
gss$predicted_class_a_i <- ifelse(gss$predicted_prob_a_i > 0.5, 1, 0)

roc_a_i <- roc(polviews ~ predicted_prob_a_i, data = gss)
roc_a_i$auc
plot(roc_a_i)

conf_matrix_a_i <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_i
```

## Confed with Age, Income, and Education

```{r, echo=TRUE}

model_a_i_e <- glm(confed ~ age + income + educ, data = gss, family = binomial)
summary(model_a_i_e)

1 - (model_a_i_e$deviance / model_a_i_e$null.deviance)

gss$predicted_prob_a_i_e <- predict(model_a_i_e, type = "response")
gss$predicted_class_a_i_e <- ifelse(gss$predicted_prob_a_i_e > 0.5, 1, 0)

roc_a_i_e <- roc(polviews ~ predicted_prob_a_i_e, data = gss)
roc_a_i_e$auc
plot(roc_a_i_e)

conf_matrix_a_i_e <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i_e),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_i_e

```

## Confed with Age, Income, Education, and Who R Voted For

```{r, echo=TRUE}

model_a_i_e_p <- glm(confed ~ age + income + educ + pres, data = gss, family = binomial)
summary(model_a_i_e_p)

1 - (model_a_i_e_p$deviance / model_a_i_e_p$null.deviance)

gss$predicted_prob_a_i_e_p <- predict(model_a_i_e_p, type = "response")
gss$predicted_class_a_i_e_p <- ifelse(gss$predicted_prob_a_i_e_p > 0.5, 1, 0)

roc_a_i_e_p <- roc(polviews ~ predicted_prob_a_i_e_p, data = gss)
roc_a_i_e_p$auc
plot(roc_a_i_e_p)

conf_matrix_a_i_e_p <- confusionMatrix(
  data = as.factor(gss$predicted_class_a_i_e_p),
  reference = as.factor(gss$polviews)
)
conf_matrix_a_i_e_p
```

## Confed with Age, Income, Education, Who R Voted For, and Year

```{r, echo=TRUE}

model_all <- glm(confed ~ age + income + educ + pres + year, data = gss, family = binomial)
summary(model_all)


1 - (model_all$deviance / model_all$null.deviance)

gss$predicted_prob_all <- predict(model_all, type = "response")
gss$predicted_class_all <- ifelse(gss$predicted_prob_all > 0.5, 1, 0)

roc_all <- roc(polviews ~ predicted_prob_all, data = gss)
roc_all$auc
plot(roc_all)

conf_matrix_all <- confusionMatrix(
  data = as.factor(gss$predicted_class_all),
  reference = as.factor(gss$polviews)
)
conf_matrix_all
```

## Confed with Age, Income, Education, and an Interaction Between Who R Voted For and Incumbency

```{r, echo=TRUE}

interaction_model_wo_year <- glm(confed ~ age + income + educ + pres * incumbency, data = gss, family = binomial)
summary(interaction_model_wo_year)

1 - (interaction_model_wo_year$deviance / interaction_model_wo_year$null.deviance)

gss$predicted_prob_wo_year <- predict(interaction_model_wo_year, type = "response")
gss$predicted_class_wo_year <- ifelse(gss$predicted_prob_wo_year > 0.5, 1, 0)

roc_wo_year <- roc(polviews ~ predicted_prob_wo_year, data = gss)
roc_wo_year$auc
plot(roc_wo_year)

conf_matrix_wo_year <- confusionMatrix(
  data = as.factor(gss$predicted_class_wo_year),
  reference = as.factor(gss$polviews)
)
conf_matrix_wo_year
```

## Confed with Age, Income, Education, Year, and an Interaction Between Who R Voted For and Incumbency

```{r, echo=TRUE}

interaction_model_year <- glm(confed ~ age + income + educ + year + pres * incumbency, data = gss, family = binomial)
summary(interaction_model_year)


1 - (interaction_model_year$deviance / interaction_model_year$null.deviance)

gss$predicted_prob_year_interaction <- predict(interaction_model_year, type = "response")
gss$predicted_class_year_interaction <- ifelse(gss$predicted_prob_year_interaction > 0.5, 1, 0)

roc_year_interaction <- roc(polviews ~ predicted_prob_year_interaction, data = gss)
roc_year_interaction$auc
plot(roc_year_interaction)

conf_matrix_year_interaction <- confusionMatrix(
  data = as.factor(gss$predicted_class_year_interaction),
  reference = as.factor(gss$polviews)
)
conf_matrix_year_interaction
```
